---
title: "class8"
author: "Amir Ferry"
date: "10/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means example

Make up some data to cluster

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3)) # rnorm will give me 30 points around -3/+3
x <- cbind(x=tmp, y=rev(tmp)) 
plot(x)
```
Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results
Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
k <- kmeans(x, 2, nstart=20)
```
```{r}
k
```
```{r}
k$size # Q. How many points are in each cluster?
k$centers # Cluster center?
k$cluster # Assignment
```
```{r}
plot(x, col = k$cluster)
points(k$centers, col = "blue", pch = 15)
```

 ## Hierarchical clustering in R
 
```{r}
hc <- hclust( dist(x) ) # Directly passing distance through hclust (instead of storing)
```
 
 Plot my results
```{r}
plot(hc)
abline(h=6, col="red")
abline(h=4, col = "blue")
cutree(hc, h = 6)
```
 
 ## Note you can see the two clusters (1-30 and 31-60) without telling it to show 2
 
```{r}
grps <- cutree(hc, h=4)
table(grps)
```
 
 I can also cut tree to a defined # of groups using k = #
```{r}
grps <- cutree(hc, k = 2)
```
 
```{r}
plot(x, col = grps)
```
 
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```
 
```{r}
plot(x)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```
```{r}
hc <- hclust(dist(x))
plot(hc)
```
 
```{r}
hc <- hclust(dist(x))
grps <- cutree(hc, k = 3)
plot(x, col = grps) # assign color according to clustering
```
 
```{r}
table(grps)
```
 
```{r}
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV", row.names=1) 

head(mydata)
```

```{r}
nrow(mydata)
ncol(mydata)
```


## NOTE: prcomp() expects the samples to be rows and
genes to be columns so we need to first transpose the
matrix with the t() function!

```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE) # Flips the matrix
```
 
```{r}
pca
pca$x[,1]
```
 
```{r}
plot(pca$x[,1], pca$x[,2])
```
 
```{r}
summary(pca)
```
 
```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per
```
 
```{r}
plot(pca)
```
 
```{r}
plot(pca$x[,1:2], col = c("red", "red", "red", "red", "red", "blue", "blue", "blue", "blue", "blue"))
```
## PCA Exercise - Europe Food Trends

```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
nrow(x)
ncol(x)
```

```{r}
head(x)
```

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x))) # beside juxtaposes or stacks data
```
```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
plot(pca$x[,1:2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1:2], colnames(x), col = (1:4))
```










