---
title: "class9"
author: "Amir Ferry"
date: "10/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
wisc.df <- read.csv("WisconsinCancer (1).csv")
```
```{r}
head(wisc.df)
```

Here we examine data from `r nrow(wisc.df)` patient samples.

```{r}
x <- table(wisc.df$diagnosis) # Use table function to tabulate # each diagnoses 
```

In this data set, we have `r x["M"]` cancer and `r x["B"]` non-cancer


```{r}
colnames(wisc.df)
```
```{r}
grep("_mean", colnames(wisc.df), value = TRUE)
```

To find out how many there are, I can call `length()` on the result of `grep()` 

```{r}
length(grep("_mean", colnames(wisc.df), value = TRUE))
```
```{r}
wisc.data <- as.matrix(wisc.df[,3:32]) # Convert to a matrix, excluding ID and X columns
```

```{r}
round(apply(wisc.data, 2, sd), 3)
```
```{r}
# Perform PCA

wisc.pr <- prcomp(wisc.data, scale = TRUE) # SD shows lots of variation so need to scale
summary(wisc.pr)
```
```{r}
# Plot using biplot function
biplot(wisc.pr)
```
```{r}
plot(wisc.pr) #Screeeee plot
```
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2],
     xlab = "PC1", ylab = "PC2", col = wisc.df$diagnosis) # Color by diagnoses
```

```{r}
x <- summary(wisc.pr)
```
```{r}
x$importance[,"PC1"]
```

The first PC captures `r x$importance[2,"PC1"] * 100 ` of the original variance 

> Q How many PC's are required to describe at least 70% of the data?

```{r}
which(x$importance[3,] > 0.7) [1] # Give me first value that captures >70%
```

Hierarchical Clustering

```{r}
data.scaled <- scale(wisc.data)
```
```{r}
data.dist <- dist(data.scaled)
```

Create hierarchical cluster model using complete linkage

```{r}
wisc.hclust <- hclust(data.dist, "complete")
```
```{r}
plot(wisc.hclust)
abline(h = 19, col = "red", lty = 2)

```

```{r}
dist.pca <- dist(wisc.pr$x[,1:7])
wisc.pr.hclust <- hclust(dist.pca, method = "ward.D2")
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
table(grps, wisc.df$diagnosis)
```
```{r}
plot(wisc.pr$x[,1:2], col = grps)
```

```{r}
g <- as.factor(grps)
levels(g)
```
```{r}
g <- relevel(g,2)
levels(g)
```

```{r}
plot(wisc.pr$x[,1:2], col=g)
```

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col = g)
points(npc[,1:2], col = "blue", pch = 16, cex = 3)
text(npc[,1], npc[,2], c(1,2), col="white")
```




















